Parallel Computations for Solving Multicriteria Mixed-Integer Optimization Problems

### 2020-12-16 2021/PСT_GergelKozinov/2021_PCT.tex
fix some misprints

In the computational scheme (\ref{eq:11}), for any value of continuous parameters $y \in D$, the values of the function $F(\alpha,y,u)$ are calculated for all possible values of discrete parameters $u \in U$. However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of discrete parameters $u^*$, so the calculation of function $F(\alpha,y,u)$ values for other values of discrete parameters is redundant. Thus, for example, for the problem
\begin{equation}\label{eq:12}
 \min\{u^2(sin(x)+sin(10x/3)):x\in[2.7,7.5],u\in \{1,2\}\}
\end{equation}
having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form from Fig.~\ref{fig:1} 
\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig1}
  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=2$)}
  \label{fig:1}
\end{figure}
and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.

@@

The final stage in improving the efficiency of solving MCOmix problems in the proposed approach is to organize parallel computing on parallel high-performance systems. Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded. For example, data parallelization (splitting the computation domain between available computating elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.  In \cite{c24,c32} a new approach to parallelizing calculations for solving global optimization problems is proposed: parallelism of calculations is provided by organizing simultaneous calculation of the values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain $D$. This approach makes it possible to parallelize the most time-consuming part of the global search process, and, due to its general nature -- it can be applied to almost any global search method for a wide variety of global optimization problems.

>>>>

In the computational scheme (\ref{eq:11}), for any value of continuous parameters $y \in D$, the values of the function $F(\alpha,y,u)$ are calculated for all possible values of discrete parameters $u \in U$. However, the smallest value of the function $F(\alpha,y,u)$ is achieved only with one specific value of discrete parameters $u^*$, so the calculation of function $F(\alpha,y,u)$ values for other values of discrete parameters is redundant. Thus, for example, for the problem
\begin{equation}\label{eq:12}
 \min\{u^2(sin(x)+sin(10x/3)):x\in[2.7,7.5],u\in \{1,2\}\}
\end{equation}
having one continuous and one discrete parameter, the plots of the  function $F(\alpha,y,u)$ for different values of the discrete parameter have form shown in Fig.~\ref{fig:1} and, as one can see, there is no need to calculate the value of the function $F(\alpha,y,u)$ for $u=1$.

\begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{fig1}
  \caption{The function $F(\alpha,y,u)$ of the problem (\ref{eq:12}) plotted for different values of the discrete parameter (left, $u=1$,  right, $u=2$)}
  \label{fig:1}
\end{figure}

@@

The final stage in improving the efficiency of solving MCOmix problems in the proposed approach is to organize parallel computing on parallel high-performance systems. Unfortunately, attempts to develop parallel variants of AGMIS and AMMIS algorithms using existing parallelization methods have not succeeded. For example, data parallelization (splitting the computation domain between available computing elements) results in only one processor processing a subdomain of the search domain  containing the sought globally optimal solution of the problem, with other processors performing redundant calculations.  In \cite{c24,c32} a new approach to parallelizing calculations for solving global optimization problems is proposed: parallelism of calculations is provided by organizing simultaneous calculation of the values of the minimized function $F(\alpha,y,u)$ from (\ref{eq:8}) at several different points of the search domain $D$. This approach makes it possible to parallelize the most time-consuming part of the global search process, and, due to its general nature -- it can be applied to almost any global search method for a wide variety of global optimization problems.
